\section{Results}

My program, \mbt, has proven to be a competent poker player when facing simple opponents. It can consistently beat \sbt in a two player match. It can also beat three separate instances of \sbt in a four player match.

I have also demonstrated the effects on performance of varying factors such as the thinking time, the presence of the opponent models, the selection strategy and the number of opponents. 

%I am very pleased with these results and with the success of the project as a whole.



\section{Future Work}

I think the biggest weaknesses of \mbt are its limited opponent models. Currently, they are too specific to \sbt and too specific to games with only two players. This means that \mbt will often make poor choices when facing other opponents or playing in games with more than two players. 

I would like to expand the opponent models for use against other opponents. I also think it would be interesting to create an opponent model with the ability to adapt to unknown opponents or maybe one which could learn from its own mistakes. 

%I also think that the way in which I implemented the opponent models is not particularly elegant. Given more time 

%I also think that there is potential to improve the simulation strategy. 



\section{Summary}

% summary of whole project
% stuck to plan?
% errors made

In this project, I have successfully built a working poker bot. I implemented the \mcts algorithm with several alternative strategies and I created two simple opponent models. 

%I stuck to my original plan quite closely although it took a lot longer than I had anticipated. If I had to restart this project, I would do two things differently. First, I would use opentestbed instead of \pa. \pa is not ideal for evaluating poker bots and it provides no way to control the random elements in poker. Second, I would approach the opponent models in a different way. I think that my current design is a little too difficult to maintain; making simple changes can sometimes require several different files to be edited.

I stuck to my original plan quite closely although it took a lot longer than I had anticipated. If I could go back and change one thing, I would have chosen opentestbed instead of \pa. \pa is not ideal for evaluating poker bots and it does not provide a way to control the random elements in poker.


However, I am delighted with the results I have achieved and with the success of the project as a whole.







